---
title: "EDA FINAL PROJ"
author: "Hermann Fan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
getwd()
```

```{r}
# load the dataset
df <- read.csv("data_tables/eu/data.csv")
head(df)
```
```{r}
# STEP 1: DATA CLEANING
df_clean <- na.omit(df)# get rid of the missing values
```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# View basic statistics
summary(df)
str(df)
```

```{r}
df %>%
  group_by(Vaccine) %>%
  summarise(TotalDoses = sum(FirstDose + SecondDose, na.rm = TRUE)) %>%
  ggplot(aes(x = Vaccine, y = TotalDoses, fill = Vaccine)) +
  geom_bar(stat = "identity") +
  ggtitle("Total Vaccine Doses by Vaccine Type") +
  xlab("Vaccine Type") +
  ylab("Total Doses") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Load necessary library
library(dplyr)

# Read the CSV files
death_df <- read.csv("data_tables/eu/death.csv")
# Aggregate data to reduce size, including positivity_rate
data_aggregated <- df %>%
  group_by(ReportingCountry) %>%
  summarise(
    Denominator = mean(Denominator, na.rm = TRUE),
    NumberDosesReceived = mean(NumberDosesReceived, na.rm = TRUE)
  )

death_aggregated <- death_df %>%
  group_by(country_code) %>%
  summarise(
    new_cases = mean(new_cases, na.rm = TRUE),
    tests_done = mean(tests_done, na.rm = TRUE),
    positivity_rate = mean(positivity_rate, na.rm = TRUE) # Include positivity rate
  )

# Merge the datasets
merged_reduced_df <- merge(data_aggregated, death_aggregated, 
                           by.x = "ReportingCountry", by.y = "country_code", all = FALSE)

# Replace NA values with the median for numerical columns
for (col in names(merged_reduced_df)) {
  if (is.numeric(merged_reduced_df[[col]])) {
    median_value <- median(merged_reduced_df[[col]], na.rm = TRUE)
    merged_reduced_df[[col]][is.na(merged_reduced_df[[col]])] <- median_value
  }
}

# View the final dataset
print(merged_reduced_df)

# Optionally, write the final dataset to a CSV file
write.csv(merged_reduced_df, "merged_reduced_data_with_positivity_rate.csv", row.names = FALSE)
```

```{r}
merged <- read.csv("data_tables/eu/merge.csv")
head(merged)

summary(merged)
colSums(is.na(merged))

pairs(merged[, sapply(merged, is.numeric)], main = "Pairwise Scatterplot")
library(corrplot)
corr_matrix <- cor(merged[, sapply(merged, is.numeric)], use = "complete.obs")
corrplot(corr_matrix, method = "circle")
```
# T TEST 1

```{r}
merged$NumberDosesReceived <- as.numeric(merged$NumberDosesReceived)
merged$new_cases <- as.numeric(merged$new_cases)

merged <- na.omit(merged)

high_doses <- subset(merged, NumberDosesReceived > median(merged$NumberDosesReceived))
low_doses <- subset(merged, NumberDosesReceived <= median(merged$NumberDosesReceived))

t_test_result <- t.test(high_doses$new_cases, low_doses$new_cases, alternative = "two.sided")

print("T-Test Result:")
print(t_test_result)
```
# T TEST 2
Countries with a higher denominator (population) have significantly different mean new cases (new_cases) compared to countries with a lower denominator.
```{r}
high_denominator <- subset(merged, Denominator > median(merged$Denominator))
low_denominator <- subset(merged, Denominator <= median(merged$Denominator))

t_test_denominator <- t.test(high_denominator$new_cases, low_denominator$new_cases, alternative = "two.sided")

print("T-Test for Denominator and New Cases:")
print(t_test_denominator)
```

```{r}
merged$new_case_standerdized <- (merged$new_cases / merged$Denominator)*100
head(merged)
```
# T Test 3 --> Does the Number Doses Received affect the new case per capita?
H0: No effect
H1: Yes there is an effect

```{r}
# Ensure columns are numeric
merged$NumberDosesReceived <- as.numeric(merged$NumberDosesReceived)
merged$new_case_standerdized <- as.numeric(merged$new_case_standerdized)

# Perform a paired t-test
t_test_result <- t.test(merged$NumberDosesReceived, merged$new_case_standerdized, paired = TRUE)

print("T-Test Result (Number of Doses Received vs New Cases):")
print(t_test_result)
```
> If we are setting the significant level as 5%, P-VALUE actually greater than the Alpha, we failed to reject H0, but it's not that far away, I will go with Ha is right.

```{r}
# Ensure columns are numeric
merged$NumberDosesReceived <- as.numeric(merged$NumberDosesReceived)
merged$positivity_rate <- as.numeric(merged$positivity_rate)

t_test_result <- t.test(merged$NumberDosesReceived, merged$positivity_rate, paired = TRUE)

print("T-Test Result (Number of Doses Received vs New Cases):")
print(t_test_result)
```

```{r}
ggplot(merged, aes(x = reorder(ReportingCountry, -positivity_rate), y = positivity_rate)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(
    title = "Positivity Rate by Country",
    x = "Country",
    y = "Positivity Rate"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
```
# New Dataset with death_rate:

```{r}
dr_df <- read.csv("data_tables/eu/d_r.csv")
head(dr_df)
```
```{r}
total_deaths_by_country <- aggregate(deaths ~ geoId, data = dr_df, sum)

ggplot(total_deaths_by_country, aes(x = reorder(geoId, -deaths), y = deaths)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    title = "Total Deaths by Country",
    x = "Country",
    y = "Total Deaths"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

```
```{r}
total_cases_by_year <- aggregate(cases ~ year, data = dr_df, sum)

ggplot(total_cases_by_year, aes(x = year, y = cases)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "darkblue", size = 3) +
  labs(
    title = "Total Cases by Year",
    x = "Year",
    y = "Total Cases"
  ) +
  theme_minimal()
```
```{r}
total_deaths_by_year <- aggregate(deaths ~ year, data = dr_df, sum)

# Create a line plot of total cases vs year
ggplot(total_deaths_by_year, aes(x = year, y = deaths)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "darkblue", size = 3) +
  labs(
    title = "Total deaths by Year",
    x = "Year",
    y = "Total Deaths"
  ) +
  theme_minimal()
```

